---
title: "Checking the quality of the Cytometry Data"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
knit: (function(inputFile, encoding) { out_dir <- '../../output/Notebook'; rmarkdown::render(inputFile,
  encoding=encoding, output_dir=file.path(dirname(inputFile), out_dir)) })
---

```{r, message=FALSE}
library(dplyr)
library(stringr)
library(ggplot2)
library(flowCore)
```


```{r}
strain_meta = read.csv('../../input/meta/Strain_Meta.csv')
row_meta = read.csv('../../input/meta/Row_Estradiol_Meta.csv')
col_meta = read.csv('../../input/meta/Column_Progesterone_Meta.csv')
meta = left_join(strain_meta,col_meta) %>% merge(row_meta)
head(meta)
```
Now lets calculate some book keeping variables:

```{r}
meta = meta %>%
    mutate(well_file = paste0('Specimen_001_',
                              Row,
                              Column,
                              '_',
                              Row,
                              str_pad(Column, 2, pad = "0"),
                              '.fcs' ),
           pathway_path = paste0('../../input/fcs/pathway/', 
                                 Strain,
                                 '/',
                                 well_file),
           survive_path = paste0('../../input/fcs/survive/', 
                                 Strain,
                                 '/',
                                 well_file))
meta$Row = factor(meta$Row, levels=rev(levels(meta$Row)))
meta$Column = factor(meta$Column)
meta$Progesterone_Factor = factor(round(meta$Progesterone_Dose, digits=1))
meta$Estradiol_Factor = factor(round(meta$Estradiol_Dose, digits=1))


meta %>% head
```
# Pathway activity orientation
Now lets write a little function to extract the median corrected FITC value for an fcs file
```{r}
get_summary_statistics = function(file_path){
    path = as.character(file_path)
    fcs <- exprs(read.FCS(path, transformation=FALSE, truncate_max_range = FALSE))
    clean = (fcs[ ,'FSC-H'] > 0) & 
            (fcs[ ,'SSC-H'] > 0) & 
            (fcs[ ,'FITC-H'] > 0) & 
            (fcs[ ,'mCherry-H'] > 0)
    fcs = fcs[clean,]
    fitc_norm = fcs[,'FITC-H'] / fcs[,'SSC-H']
    out = median(log10(fitc_norm))
    return(out)
}

meta = meta %>% 
    mutate(fitc = Vectorize(get_summary_statistics)(pathway_path) + 1) %>%
    group_by(Strain) %>% mutate(fitc_norm = fitc / fitc[Estradiol_Dose == 0 & Progesterone_Dose == 0] ) %>% ungroup()
```

Now lets plot the plates to see their structure:

```{r}
meta %>% ggplot(aes(x=Column,y=Row, fill = fitc_norm)) + geom_tile() + facet_grid(Repressor~Activator)
```

It is clear that the activity of the pathway goes up when we traverse the plate from right to left (activator dimension). 
It is also clear that the activity of the pathway goes down when traverse the plate from bottom to top (repressor dimension)
this is not typical of how Andres and I do dose responses, however it may be how Elham did it.

One thing I can see from this data is that this patturn is consistant! which is a relief!

# Survival plate

First thing to do is to establish the correct cutoff. I'll first do this by just plotting. Maybe a mixed gausian mixture model might work as well.

```{r}
get_fcs = function(path){
    fcs <- exprs(read.FCS(path, transformation=FALSE, truncate_max_range = FALSE))
    return(fcs)
}

fcs = lapply(meta$survive_path, get_fcs)
fcs = do.call("rbind", fcs)
hist(log10(fcs[,'mCherry-H']))
```
It seems that 10^3 is a perfectly reasonable cutoff between dead and alive cells
Lets quatify the survival for all the plates.

```{r}
percent_alive = function(path, lim){
    fcs <- exprs(read.FCS(path, transformation=FALSE, truncate_max_range = FALSE))
    mcherry = fcs[,"mCherry-H"]
    return(sum(mcherry<lim)/length(mcherry))
}
meta = meta %>% mutate(percent_alive = Vectorize(percent_alive)(survive_path, 10^3))
```

Lets take a look

```{r}
meta %>% ggplot(aes(x=Column,y=Row, fill = percent_alive)) + geom_tile() + facet_grid(Repressor~Activator)
```

Looks just like opposite of the activity. So survivability negativly corrolates with pathway output?

```{r}
meta %>% ggplot(aes(x=fitc_norm, y=percent_alive, color = Activator)) + geom_point()
```

It also seems that MKK1 is far more toxic at high levels than RHO1, lets also check for the repressors:

```{r}
meta %>% ggplot(aes(x=fitc_norm, y=percent_alive, color = Repressor)) + geom_point()
```

It seems like MSN2 is a very powerful repressor of pathway output, and therefore resuces a large chunk of cell death.

# Recreating Raj's Plots
```{r}
meta %>% ggplot(
        aes(x = Progesterone_Factor, 
            y=fitc_norm, 
            color = Estradiol_Factor, 
            group = Estradiol_Factor)) + 
    geom_line() + 
    facet_grid(Repressor~Activator) + 
    xlab('Progesteron (nM)') + 
    ylab('Pathway Activity') + 
    theme(axis.text.x = element_text(face="bold",size=4, angle=270)) + 
    guides(color=guide_legend(title='Estradiol (nM)'))

```